\hspace{1cm} Do syntactic and semantic similarity lead to interference effects? \newline Evidence from self-paced reading and event-related potentials using German

### 2023-12-31 paper/main.tex
removed paragraph

An obvious difference from previous studies is the experimental methodology. The majority of the previous experiments employed eye-tracking while reading, while we used a self-paced reading task. Self-paced reading (SPR) could trigger a greater tendency to cause interference effects compared to natural reading (as in eye-tracking studies), due to the restricted reading format in SPR: Comprehenders have no opportunity to revisit previous material during conventional self-paced reading \citep[but see][]{BSPR}. SPR is likely to be more demanding on the comprehender's memory, which might make it easier to detect interference effects. This argumentation is supported by our estimation of the results of \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1, the only other SPR experiment using this design: %(see Figure \ref{fig:previous} in the appendix):
In the critical region, the estimates of both main effects (syntactic and semantic interference) are very similar to the results of the present study; the interaction effect observed by \citeauthor{vandyke07} in the spill-over region is also similar to our estimate. However, it remains unclear why semantic interference was especially prominent in our study compared to the previous ones.

There is one further possible reason for the diverging effects in our studies compared to previous work: the theoretically-predicted effects we are investigating here may be too subtle to consistently be detectable across studies. In several other attempts to replicate previously observed findings, we have often found it impossible to replicate the originally claimed effects, despite spending four or more years collecting data to obtain sufficient evidence. For example, contrary to published claims \parencite[e.g.,][]{vandyke_mcelree06}, in a large cross-linguistic study, we found evidence against proactive interference from sentence-external distractors \parencite{mertzen_repVDM2006}. A second example is the well-known syntactic local coherence effect in English \parencite{taboretal04}; in a recent large-sample study, we have found evidence against syntactic local coherence \parencite{lcpaape2023}.  A third example is the interaction of locality and surprisal effects reported in \textcite{levy&keller_2013}; we attempted and failed to replicate the key claims in seven experiments \parencite{vasishth2018_signficancefilter}. A fourth example is the processing of antecedent-reflexive dependencies; \textcite{dillon2013} argued that reflexives are exempt from interference, but a large-scale direct replication attempt  \parencite{jaeger_etal_2020} found evidence that reflexives do show interference effects \parencite[cf.][]{yadav2021individual}. Finally, we have also failed to replicate the surprising claim  \parencite{gibsonwu,hsiao03} that Chinese object relatives are easier to process than subject relatives \parencite{VasishthetalPLoSOne2013,VasishthetalPLoSOne2013}. It is in principle possible that all the originally claimed effects actually represent true effects but, due to the inherent variability of data, the effects may be impossible to detect consistently in replications.  Independent of any power issues in all these different studies, it is possible that it is unreasonable to expect consistent replications of effects in behavioral data, especially when investigating subtle effects.  Certain other effects in psycholinguistics, such as strong garden-path effects \parencite{frazierrayner82}, are easily and consistently replicable \parencite[e.g.,][]{huang2023surprisal}. By contrast, subtle manipulations such as interference effects and local coherence may not show up consistently in replication attempts. One strong possibility is that there are systematic and important individual differences between participants' behavior; an increased focus away from investigating average effects and toward individual differences may reveal important insights about the conditions under which subtle effects are replicable; we demonstrate this point for the antecedent-reflexive dependency in \parencite{yadav2021individual}.

Nevertheless, the contrast between our results and those from  \citeauthor{mertzen}'s (\citeyear{mertzen}) German experiment is surprising given that both used German items and partially even identical items. So, why did we find a semantic interference effect in all regions, whereas \citet{mertzen} only found one in the pre-critical region? Apart from the methodological difference (eye-tracking vs.\ self-paced reading), a plausible  explanation for the divergence is lower power in \citeauthor{mertzen}'s (\citeyear{mertzen}) study. While they already had considerably more participants and items than many psycholinguistic studies, in the present study the number of participants was much higher than in theirs (774 vs.\ 121) and the number of items was also higher (60/120 vs.\ 40). 
We will revisit the absence of syntactic interference in our data in the general discussion, where we investigate the question using computational modeling.

>>>>

An obvious difference from previous studies is the experimental methodology. The majority of the previous experiments employed eye-tracking while reading, while we used a self-paced reading task. Self-paced reading (SPR) could trigger a greater tendency to cause interference effects compared to natural reading (as in eye-tracking studies), due to the restricted reading format in SPR: Comprehenders have no opportunity to revisit previous material during conventional self-paced reading \citep[but see][]{BSPR}. SPR is likely to be more demanding on the comprehender's memory, which might make it easier to detect interference effects. This argumentation is supported by our estimation of the results of \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1, the only other SPR experiment using this design: 
In the critical region, the estimates of both main effects (syntactic and semantic interference) are very similar to the results of the present study; the interaction effect observed by \citeauthor{vandyke07} in the spill-over region is also similar to our estimate. However, it remains unclear why semantic interference was especially prominent in our study compared to the previous ones.

The contrast between our results and those from  \citeauthor{mertzen}'s (\citeyear{mertzen}) German experiment is surprising given that both used German items and partially even identical items. So, why did we find a semantic interference effect in all regions, whereas \citet{mertzen} only found one in the pre-critical region? Apart from the methodological difference (eye-tracking vs.\ self-paced reading), a plausible  explanation for the divergence is lower power in \citeauthor{mertzen}'s (\citeyear{mertzen}) study. While they already had considerably more participants and items than many psycholinguistic studies, in the present study the number of participants was much higher than in theirs (774 vs.\ 121) and the number of items was also higher (60/120 vs.\ 40). 
We will revisit the absence of syntactic interference in our data in the general discussion, where we investigate the question using computational modeling.
